# (APPENDIX) Appendix {-}

# Code outputs

## Mastering {htmltools} {#mastering-htmltools-appendix}

### Shiny RPG {#mastering-htmltools-appendix-1}
Final code expected in section \@ref(htmltools-shinyRPG):

```{r, eval = FALSE}
rpgSelect <- function(inputId, label, choices, selected = NULL,
                      multiple = FALSE, size = NULL) {
  selectTag <- shiny::selectInput(
    inputId,
    label,
    choices,
    selected,
    multiple,
    selectize = FALSE,
    width = NULL,
    size
  )

  # Modify tag
  selectTag$attribs$class <- NULL
  # Clean extra label class
  selectTag$children[[1]]$attribs$class <- NULL
  # Remove extra outer div
  selectTag$children[[2]] <- selectTag$children[[2]]$children[[1]]

  # Add good class for rppgui binding
  selectTag$children[[2]]$attribs$class <- if (is.null(size)) {
    "rpgui-dropdown"
  } else {
    "rpgui-list"
  }

  selectTag
}
```


### Shiny RPG rework {#mastering-htmltools-appendix-2}
Final code expected in section \@ref(htmltools-shinyRPG-rework):

```{r}
rpgSelect <- function(inputId, label, choices, selected = NULL,
                      multiple = FALSE, size = NULL) {
  selectTag <- shiny::selectInput(
    inputId,
    label,
    choices,
    selected,
    multiple,
    selectize = FALSE,
    width = NULL,
    size
  )
  
  selectClass <- if (is.null(size)) { # add class
    "rpgui-dropdown"
  } else {
    "rpgui-list"
  }
  
  tagQuery(selectTag)$
    removeAttrs("class")$ # remove outer div class
    find(".control-label")$
    removeAttrs("class")$ # remove class from child label
    siblings()$ # go down to the div 
    children()$ # go down to the select tag
    addClass(selectClass)$ # add class to child select
    resetSelected()$# go back to div parent
    each(function(x, i) {
      x$children[[2]] <- x$children[[2]]$children
    })$ # replace div parent 
    allTags()
}
```

## Beautify with {bslib}

### Guided exercise: Bootstrap accordions {#bslib-bootstrap-accordion-appendix}
This part provide solution for section \@ref(bslib-bootstrap-accordion).

Code for the `bs_accordion()` function is:

```{r, eval=FALSE}
bs_accordion <- function(id, items) {

  accordion_tag <- tags$div(
    class = "accordion",
    id = id,
    items
  )

  tagAddRenderHook(accordion_tag, function(x) {
    # get theme and version
    theme <- bslib::bs_current_theme()
    version <- if (bslib::is_bs_theme(theme)) {
      bslib::theme_version(theme)
    }

    if (version == "3") {
      stop(
        sprintf(
          "accordion is not available for Bootstrap %s",
          version
        )
      )
    }

    # process accordion items to add
    # missing attributes
    new_items <- lapply(seq_along(items), function(i) {

      # temp ids based on the parent id
      heading_id <- paste(id, "heading", i, sep = "_")
      controls_id <- paste0(id, "_collapse_", i)
      target_id <- paste0("#", controls_id)

      # resolve bs_according_item
      items[[i]] <- as.tags(items[[i]])

      # BS4 and BS5 have minor differences
      switch(
        version,
        "4" = tagQuery(items[[i]])$
          find(".card-header")$
          addAttrs("id" = heading_id)$
          find(".btn")$
          addAttrs(
            "data-target" = target_id,
            "aria-controls" = controls_id
          )$
          resetSelected()$
          find(".collapse")$
          addAttrs(
            "id" = controls_id,
            "aria-labelledby" = heading_id,
            "data-parent" = paste0("#", id)
          )$
          allTags(),
        "5" = tagQuery(items[[i]])$
          find(".accordion-header")$
          addAttrs("id" = heading_id)$
          children()$
          addAttrs(
            "data-bs-target" = target_id,
            "aria-controls" = controls_id
          )$
          resetSelected()$
          find(".accordion-collapse")$
          addAttrs(
            "id" = controls_id,
            "aria-labelledby" = heading_id,
            "data-bs-parent" = paste0("#", id)
          )$
          allTags()
      )
    })

    # alter main tag structure
    tagQuery(x)$
      # replace accordion items processed above
      empty()$
      append(new_items)$
      allTags()
  })
}
```

Code for the `bs_accordion_item()` function is:

```{r, eval=FALSE}
bs_accordion_item <- function(title, content, active = FALSE) {

  item_body <- tags$div(
    # id will be added from bs_accordion
    # aria-labelledby also added from bs_accordion
    # class differs between BS4 and BS5
    # data parent differs between BS4 and BS5
    class = paste("collapse", if (active) "show"),
    tags$div(
      # class differs between BS4 and BS5
      content
    )
  )

  # accordion item wrapper
  accordion_item_tag <- tags$div(
    # class differs between BS4 and BS5
    item_body
  )

  tagAddRenderHook(accordion_item_tag, function(x) {
    # get theme and version
    theme <- bslib::bs_current_theme()
    version <- if (bslib::is_bs_theme(theme)) {
      bslib::theme_version(theme)
    }

    # create accordion item header
    item_header <- if (version == "4") {
      tags$div(
        class = "card-header",
        # id will be added from bs_accordion
        tags$h2(
          class = "mb-0",
          tags$button(
            class = "btn btn-link btn-block text-left",
            type = "button",
            `data-toggle` = "collapse",
            # data-target will be added from bs_accordion
            `aria-expanded` = tolower(active),
            # aria-controls will be added from bs_accordion
            title
          )
        )
      )
    } else if (version == "5") {
      tags$h2(
        class = "accordion-header",
        tags$button(
          class = "accordion-button",
          type = "button",
          `data-bs-toggle` = "collapse",
          `aria-expanded` = tolower(active),
          title
        )
      )
    }

    # alter tag structure
    switch(
      version,
      # don't need to handle BS3
      "4" =  tagQuery(x)$
        addClass("card")$
        # prepend header tag
        prepend(item_header)$
        find(".collapse")$
        children()$
        # add class to item body
        addClass("card-body")$
        allTags(),
      "5" = tagQuery(x)$
        addClass("accordion-item")$
        prepend(item_header)$
        find(".collapse")$
        addClass("accordion-collapse")$
        children()$
        addClass("accordion-body")$
        allTags()
    )
  })
}
```

## R and JS {#from-R-to-JS-appendix}

Below is the whole websocket code obtained after following \@ref(from-R-to-JS).

```{r, eval=FALSE}
ws_handler <- function(ws) {
  # The ws object is a WebSocket object
  cat("New connection opened.\n")
  # Capture client messages
  ws$onMessage(function(binary, message) {
    # create plot
    input_message <- jsonlite::fromJSON(message)
    print(input_message)
    cat("Number of bins:", input_message$value, "\n")
    hist(rnorm(input_message$value))
    
    # update gauge widget
    output_message <- jsonlite::toJSON(
      list(
        val = sample(0:100, 1),
        message = "Thanks client! I updated the plot..."
      ),
      pretty = TRUE,
      auto_unbox = TRUE
    )
    ws$send(output_message)
    cat(output_message)
  })
  ws$onClose(function() {
    cat("Server connection closed.\n")
  })
}
```

## Shiny's input system

### Update a binding from the client {#update-binding-client-appendix}
This is the JS code from section \@ref(update-binding-client).

```js
$(function() {
  $(document).on('shiny:connected', function(event) {
    Shiny.unbindAll();
    $.extend(Shiny
      .inputBindings
      .bindingNames['shiny.actionButtonInput']
      .binding, {
        reset: function(el) {
         $(el).data('val', 0);
        },
        subscribe: function(el, callback) {
          $(el).on(
            'click.actionButtonInputBinding', function(e) {
              let $el = $(this);
              let val = $el.data('val') || 0;
              $el.data('val', val + 1);

              callback();
          });
                
          $(el).on(
            'change.actionButtonInputBinding', function(e) {
              callback();
          });
              
        }
      });
    Shiny.bindAll();
  });
      
  $('#button1').on('click', function() {
    let $obj = $('#button2');
    let inputBinding = $obj.data('shiny-input-binding');
    let val = $obj.data('val') || 0;
    inputBinding.setValue($obj, val + 10);
    $obj.trigger('change'); 
  });
  
  $('#reset').on('click', function() {
    let $obj = $('#button2');
    let inputBinding = $obj.data('shiny-input-binding');
    inputBinding.reset($obj);
    $obj.trigger('change');
  });
});
```

### Boxes on steroids {#boxes-on-steroids-appendix}
The box input binding defined in \@ref(boxes-on-steroids) may be found below.

```js
let boxBinding = new Shiny.InputBinding();
$.extend(boxBinding, {
  find: function(scope) {
    return $(scope).find('.box');
  },
  getValue: function(el) {
    let isCollapsed = $(el).hasClass('collapsed-box')
    return {collapsed: isCollapsed}; // this will be a list in R
  },
  setValue: function(el, value) {
    $(el).toggleBox();
  }, 
  receiveMessage: function(el, data) {
    this.setValue(el, data);
    $(el).trigger('change');
  },
  subscribe: function(el, callback) {
    $(el).on(
      'click', 
      '[data-widget="collapse"]', 
      function(event) {
        setTimeout(function() {
          callback();
        }, 50);
    }); 
    
    $(el).on('change', function(event) {
      setTimeout(function() {
        callback();
      }, 50);
    });
  },
  unsubscribe: function(el) {
    $(el).off('.boxBinding');
  }
});

Shiny.inputBindings.register(boxBinding, 'box-input');

$(function() {
  // overwrite box animation speed. 
  // Putting 500 ms add unnecessary delay for Shiny.
  $.AdminLTE.boxWidget.animationSpeed = 10;
});
```

### Going further {#boxes-on-steroids-more-appendix}
Final code from section \@ref(boxes-on-steroids) is given below.

```js
let boxBindingEnchanced = new Shiny.InputBinding();
$.extend(boxBindingEnchanced, {
  // Other methods from boxBinding.js: find, getValue, ...
  //...
  
  // user defined binding: extract config script
  _getConfigScript: function(el) {
    return(
      $(el)
        .parent()
        .find("script[data-for='" + el.id + "']")
    )
  },
  // user defined binding: process config script
  _processConfig: function(el) {
    return(
      JSON.parse(
        this
        ._getConfigScript(el)
        .html()
      )
    )
  },
  // user defined binding: update box width
  _updateWidth: function(el, o, n) {
    // removes old class
    $(el).parent().toggleClass("col-sm-" + o);
    $(el).parent().addClass("col-sm-" + n); 
    // trigger resize so that output resize
    $(el).trigger('resize');
  },
  // Input binding default method 
  setValue: function(el, value) {
    let config = this._processConfig(el);
    
    // JS logic
    if (value.action === "update") {
      if (value.options.hasOwnProperty("width")) {
        if (value.options.width !== config.width) {
          this._updateWidth(
            el, 
            config.width, 
            value.options.width
          )
          config.width = value.options.width;
        }
      }
      // other items to update
      
      // replace the old JSON config by the new one 
      // to update the input value 
      this
        ._getConfigScript(el)
        .replaceWith(
          '<script type="application/json" data-for="' + 
          el.id + 
          '">' + 
          JSON.stringify(config) + 
          '</script>'
        );
      
    } else if (action === "toggle") {
      $(el).toggleBox();
    }
    
  }
});

Shiny.inputBindings.register(boxBindingEnchanced, 'box-input-enhanced');

$(function() {
  // overwrite box animation speed. 
  // Putting 500 ms add unnecessary delay for Shiny.
  $.AdminLTE.boxWidget.animationSpeed = 10;
});
```

## Dynamically manage content with handlers {#shiny-custom-handlers-appendix}

### A chat system for {shinydashboardPlus} {#custom-handlers-chat-appendix}

Final code obtained in \@ref(custom-handlers-chat) is shown below.

```js
// This code creates a custom handler for userMessages
Shiny.addCustomMessageHandler(
  "user-messages", function(message) {
  let id = message.id, 
    action = message.action, 
    content = message.body, 
    index = message.index;
  
  // message text
  // We use Shiny.renderHtml to handle the case where 
  // the user pass input/outputs in the updated content 
  // that require a new dependency not available in the 
  // page at startup. 
  if (content.hasOwnProperty("text")) {
    let text;
    if (content.text.html === undefined) {
      text = content.text;
    } else {
      text = Shiny.renderHtml(
        content.text.html, 
        $([]), 
        content.text.dependencies
      ).html;
    } 
  }
  
  // unbind all
  Shiny.unbindAll();
  
  if (action === "remove") {
    $("#" + id)
      .find(".direct-chat-msg")
      .eq(index - 1)
      .remove();
  } else if (action === "add") {
    let author = content.author, 
      date = content.date, 
      image = content.image, 
      type = content.type;
    
    // build the new message 
    let newMessage = `
      <div class="direct-chat-info clearfix">
        <span class="direct-chat-name"> 
          ${author}
        </span>
        <span class="direct-chat-timestamp" 
          style="margin-left: 4px">
          ${date}
        </span>
      </div>
      <img class="direct-chat-img" src="${image}"/> 
      <div class="direct-chat-text">${text}</div>`;
      
    // build wrapper
    let newMessageWrapper;
    if (type === "sent") {
      newMessageWrapper = `
        <div class="direct-chat-msg right">
          ${newMessage} 
        </div>`;
    } else {
      newMessageWrapper = `
        <div class="direct-chat-msg">
          ${newMessage} 
        </div>`;
    }
    
    // append message
    $("#" + id)
      .find(".direct-chat-messages")
      .append(newMessageWrapper);
  } else if (action === "update") {
    
    // today's date
    let d = new Date();
    let month = d.getMonth() + 1;
    let day = d.getDate();
    let today = d.getFullYear() + '/' +
      ((''+month).length<2 ? '0' : '') + month + '/' +
      ((''+day).length<2 ? '0' : '') + day;
      
    // we assume only text may be updated. 
    // Does not make sense to modify author/date
    
    $("#" + id)
      .find(".direct-chat-text")
      .eq(index - 1)
      .replaceWith(`
        <div class="direct-chat-text">
          <small class="text-red">
            (modified: ${today})
          </small>
          <br>
        </div>
      `)
  }
  
  // Calls .initialize() for all of the input 
  // objects in all input bindings,
  // in the given scope (document)
  Shiny.initializeInputs();
  Shiny.bindAll(); // bind all inputs/outputs
});
```

## Case Study 2: Mobile development with Shiny

### Reconstruct {shinyMobile} {#mobile-shinyMobile-appendix}

The `init.js` code obtained in \@ref(mobile-shinyMobile) is shown below. 

```js
$( document ).ready(function() {
  // collect all data elements stored in body
  let config = $(document).find("script[data-for='app']");
  config = JSON.parse(config.html());

  // always erase existing root value just in case 
  // the user changes the root. This may be harmful
  config.root = "#app";
  
  // store app methods
  config.methods =  {
      toggleDarkTheme: function() {
        var self = this;
        var $html = self.$("html");
        $html.toggleClass("theme-dark");
      }
    };

  // create app instance
  app = new Framework7(config);

  // init main view
  let mainView = app.views.create('.view-main');
  
  // tapHold custom css
  if (config.hasOwnProperty("touch")) {
    if (config.touch.tapHold) {
      $("<style>")
        .prop("type", "text/css")
        .html(
          `-moz-user-select: none;
          -webkit-user-select: none;
          user-select: none;`
        )
        .appendTo("head");
    }
  }

  let notification = app.notification.create({
    text: 'Hello, how are you?',
    on: {
      opened: function () {
        console.log('Notification opened');
      }
    }
  }).open();

  let otherMessage = app.notification.create({
    text: 'You look great!'
  });

  // equivalent to setTimeout ...
  app.utils.nextTick(function() {
    otherMessage.open();
  }, 1000);

  // taphold test
  $('#mybutton').on('taphold', function () {
    app.dialog.alert('Tap hold fired!');
  });

  // Set color theme
  if (config.hasOwnProperty('color')) {
    let colorCSS = app.utils.colorThemeCSSProperties(config.color);
    $('<style>')
      .prop('type', 'text/css')
      .html(`:root {
        --f7-theme-color: ${colorCSS["--f7-theme-color"]};
        --f7-theme-color-rgb: ${colorCSS["--f7-theme-color-rgb"]};
        --f7-theme-color-shade: ${colorCSS["--f7-theme-color-shade"]};
        --f7-theme-color-tint: ${colorCSS["--f7-theme-color-tint"]};
      }`)
      .appendTo("head");
  }

  // Filled theme
  if (!config.hasOwnProperty('filled')) config.filled = false;
  if (config.filled) {
    let filledCSS = `
      :root,
      :root.theme-dark,
      :root .theme-dark {
        --f7-bars-bg-color: var(--f7-theme-color);
        --f7-bars-bg-color-rgb: var(--f7-theme-color-rgb);
        --f7-bars-translucent-opacity: 0.9;
        --f7-bars-text-color: #fff;
        --f7-bars-link-color: #fff;
        --f7-navbar-subtitle-text-color: rgba(255,255,255,0.85);
        --f7-bars-border-color: transparent;
        --f7-tabbar-link-active-color: #fff;
        --f7-tabbar-link-inactive-color: rgba(255,255,255,0.54);
        --f7-sheet-border-color: transparent;
        --f7-tabbar-link-active-border-color: #fff;
      }
      .appbar,
      .navbar,
      .toolbar,
      .subnavbar,
      .calendar-header,
      .calendar-footer {
        --f7-touch-ripple-color: var(--f7-touch-ripple-white);
        --f7-link-highlight-color: var(--f7-link-highlight-white);
        --f7-button-text-color: #fff;
        --f7-button-pressed-bg-color: rgba(255,255,255,0.1);
      }
      .navbar-large-transparent,
      .navbar-large.navbar-transparent {
        --f7-navbar-large-title-text-color: #000;
      
        --r: 0;
        --g: 122;
        --b: 255;
        --progress: var(--f7-navbar-large-collapse-progress);
        --f7-bars-link-color: rgb(
          calc(var(--r) + (255 - var(--r)) * var(--progress)),
          calc(var(--g) + (255 - var(--g)) * var(--progress)),
          calc(var(--b) + (255 - var(--b)) * var(--progress))
        );
      }
      .theme-dark .navbar-large-transparent,
      .theme-dark .navbar-large.navbar-transparent {
        --f7-navbar-large-title-text-color: #fff;
    }`;
    
    $('<style>')
      .prop('type', 'text/css')
      .html(`${filledCSS}`)
      .appendTo("head");
  }

  // dark mode
  if (!config.hasOwnProperty('dark')) config.dark = false;
  if (config.dark) {
    app.methods.toggleDarkTheme();
  }
});
```

## R + Shiny + React: welcome `{reactR}` 

### Introduction to `{reactR}`

#### Exercise 2 {#reactR-exercise-2-appendix}

JSX code from section \@ref(reactR-exercise-2) may be found below.

```jsx
import { reactShinyInput } from 'reactR';
// reactstrap components
import { Button } from "reactstrap";

function ActionButton({configuration, value, setValue}) {
  let iconTag, btnCl, innerTag;
  if (...) {
    btnCl = "btn-icon";
    innerTag = <>
      <span className="btn-inner--icon">
        <i className=... />
      </span>
      <span className="btn-inner--text">...</span>
    </>;
  } else {
    innerTag = configuration.label;
  }

  let outlined;
  if (...) {
    outlined = true;
  }

  return (
    <Button
      color={configuration.status}
      className=...
      outline=...
      size=...
      onClick={() => setValue(value + 1)}>
      {innerTag}
    </Button>
  );
}

reactShinyInput(
  '.action_button', 
  'reactstrapTest.action_button', 
  ActionButton
);
```