# Other tools {#beautify-other-tools}

## CSS
Sometimes, it is tempting to write inline CSS for a tag in the `style` attribute, for instance:

```{r, eval=FALSE}
title <- h4(
  "A title",
  style = "color: cyan; text-align: center; margin-top: 100px;"
)
shinyApp(ui = title, server = function(input, output) {})
```

As the number of CSS properties grows, the code may become hard to read. The [{cascadess}](https://github.com/nteetor/cascadess) package developed by [Nathan Teetor]() provides a more readable syntax, which works well with `%>%`.

```{r, eval=FALSE}
ui <- list(
  cascadess(),
  h4(
    "A title",
    .style %>%
      text(color = "cyan", align = "center") %>%
      margin(top = 5)
  )
)
shinyApp(ui, server = function(input, output) {})
```

{cascadess} is a pleasing way to learn CSS through R. For instance, the CSS [text formating](https://www.google.com/search?client=safari&rls=en&q=text+properties+css&ie=UTF-8&oe=UTF-8) includes properties like `text-color`, `text-align`, ...The `text` function has the same parameters, namely color, align, as shown in the above example. 

Notice that we included the `cascadess` function, which is nothing more than an htmldependency object necessary to apply styles on elements.

```{r}
cascadess()
```

Try to run the previous example without this element. Nothing will be properly displayed!