# Testing and validating templates elements {#custom-templates-testing}
Until now, we have been building the template boilerplates, that is the main skeleton functions (page, navbar, navbar menu, ...) as well as some components such as cards, ribbons, progress bars. We also exposed some techniques to substantially give more interactivity to the template, leveraging our freshly adquired JavaScript skills. Does this mean we are ready to make the template public? Not yet since some essentials steps are missing:


  - Input validation is a crucial step toward success. Briefly, it consists in checking user inputs so that your functions fail safely and elegantly by providing meaningful error messages or warnings. This concept has already been covered in R for Data Science, Advanced R and a lot of other resources. Hence, I am not trying to reinvent the wheel and we will rely on already existing patterns, whose effectiveness is no longer to be demonstrated. Welcome to the defensive programming world! 
  - On the other hand, testing components allows to check if a functions does what it is supposed to do. Consequently, it is a proof of robustness, and increases reproducibility. It significantly reduces the mental load when it comes to start code refactoring, thereby making you feel slightly less guilty about creating breaking changes since, most of the time, your tests will be able to capture those error. 



## Validate template functions
Below, we will show some examples to validate user inputs. We first consider the `tabler_card` element from Chapter \@ref(custom-templates-skeleton).

```{r}
tabler_card <- function(..., title = NULL, status = NULL, width = 6, stacked = FALSE, padding = NULL) {
  
  card_cl <- paste0(
    "card", 
    if (stacked) " card-stacked",
    if (!is.null(padding)) paste0(" card-", padding)
  )
  
  div(
    class = paste0("col-md-", width),
    div(
      class = card_cl,
      if (!is.null(status)) {
        div(class = paste0("card-status-top bg-", status))
      },
      div(
        class = "card-body",
        # we could have a smaller title like h4 or h5...
        if (!is.null(title)) {
          h3(class = "card-title", title)
        },
        ...
      )
    )
  )
  
}
```

The first thing is to think about what to validate. Here, I see at least 3 test:

  - checking the status value
  - checking the width value
  - checking the padding value
  
Writing validation tests requires knowledge about the underlying mechanisms. In our case, we must know the Bootstrap 4 grid rules and valid color statuses. The Bootstrap 4 [grid](https://getbootstrap.com/docs/4.0/layout/grid/) ranges from 1 to 12. In other words, a card having a width of 12 will take the full page wherease, we may align 3 cards or width 4. Valid [statuses](https://getbootstrap.com/docs/4.0/utilities/colors/) are primary, secondary, success, info, danger, warning, light and dark.

It is therefore pretty straightforward to validate the card width: it must be numeric and between 1 and 12. Moreover, since the template has other containers including the width parameter, we will create a function

```{r}
validate_width <- function(width) {
  if (is.numeric(width)) {
    if (width < 1 || width > 12) {
      stop("width must belong to [1, 12], as per Bootstrap 4 grid documentation. See more at https://getbootstrap.com/docs/4.0/layout/grid/")
    }
  } else {
    stop("width must be numeric")
  }
}

# validate_width(-1)
# validate_width(13)
# validate_width("string")
```

To check the status parameter, we save the valid statuses in a vector and take the `validStatuses` function from shinydashboard [utils](https://github.com/rstudio/shinydashboard/blob/master/R/utils.R). It also make sense to create a function since this parameter is widely used among template components. Contrary to the shinydashboard function, our custom `valid_status` does not fail if the status is NULL since it is not a mandatory parameter.

```{r}
valid_statuses <- c(
  "primary", 
  "secondary",
  "success", 
  "info", 
  "warning", 
  "danger",
  "light",
  "dark"
)

validate_status <- function(status) {

  if (is.null(status)) {
    return(TRUE) 
  } else {
    if (status %in% valid_statuses) {
      return(TRUE)
    }
  }

  stop("Invalid status: ", status, ". Valid statuses are: ",
       paste(valid_statuses, collapse = ", "), ".")
}
```

We finish by the padding validation. According to the tabler documentation, there are three possible choices, namely `sm`, `md` or `lg`.

```{r}

valid_paddings <- c("sm", "md", "lg")

validate_padding <- function(padding) {
  if (!is.null(padding)) {
    if (!(padding %in% valid_paddings)) {
      stop("Invalid status: ", padding, ". Valid choices are: ",
       paste(valid_paddings, collapse = ", "), ".")
    }
  }
}
```

We apply all these validation to our card element.

```{r}
tabler_card <- function(..., title = NULL, status = NULL, width = 6, stacked = FALSE, padding = NULL) {
  
  validate_status(status)
  validate_width(width)
  validate_padding(padding)
  
  card_cl <- paste0(
    "card", 
    if (stacked) " card-stacked",
    if (!is.null(padding)) paste0(" card-", padding)
  )
  
  div(
    class = paste0("col-md-", width),
    div(
      class = card_cl,
      if (!is.null(status)) {
        div(class = paste0("card-status-top bg-", status))
      },
      div(
        class = "card-body",
        # we could have a smaller title like h4 or h5...
        if (!is.null(title)) {
          h3(class = "card-title", title)
        },
        ...
      )
    )
  )
  
}
```

As of R 4.0.0, the `stopifnot` function may be a good alternative to `stop`. 

We recommand to be reasonable and not to validate every single parameter.


## Testing templates elements
Imagine if one day, someone or yourself accidentaly remove the width validation function, after a significant refactoring. Later, you receive a new message from github, stating that someome opened a new issue. According to the report, the card is not properly displayed altough no error is thrown. Among the 400 lines of code provided, you hardly notice that the width parameter is set to 13, which is not in line with the Bootstrap 4 documentation, as it should remain between 1 and 12. You lost 10 minutes, so is your end user, which is even worse. 

With a proper testing pipeline, this problem could have been avoided. Ironically, writing the corresponding test takes only 2 minutes.

A reference for testing functions is the [{testthat}](https://testthat.r-lib.org/index.html) package. In short, a unit test consists in setting expectations about our function and check weither they fail or pass. For instance, with our `tabler_card` example, the `validate_width` must fail if the given width is not in the expected bounds or not numeric. We apply the `test_that` function with a description of the test context, followed by the expectations inside the curly brackets.

```{r}
test_that("validate width works", {
  expect_error(tabler_card(width = -1))
  expect_error(tabler_card(width = 13))
  expect_error(tabler_card(width = "hello world"))
})
```

We then test `validate_status` and `validate_padding`.

```{r}
test_that("validate status works", {
  expect_error(tabler_card(status = "toto"))
})

test_that("validate padding works", {
  expect_error(tabler_card(width = "xs"))
})
```

So far so good. In few lines of code, we substantially increased the robustness of our function without increasing its complexity. Now, let's try to remove the `validate_width` step from the `tabler_card`.

```{r, eval=FALSE}
tabler_card <- function(..., title = NULL, status = NULL, width = 6, stacked = FALSE, padding = NULL) {
  
  validate_status(status)
  validate_padding(padding)
  
  card_cl <- paste0(
    "card", 
    if (stacked) " card-stacked",
    if (!is.null(padding)) paste0(" card-", padding)
  )
  
  div(
    class = paste0("col-md-", width),
    div(
      class = card_cl,
      if (!is.null(status)) {
        div(class = paste0("card-status-top bg-", status))
      },
      div(
        class = "card-body",
        # we could have a smaller title like h4 or h5...
        if (!is.null(title)) {
          h3(class = "card-title", title)
        },
        ...
      )
    )
  )
  
}


test_that("validate width works", {
  expect_error(tabler_card(width = -1))
  expect_error(tabler_card(width = 13))
  expect_error(tabler_card(width = "hello world"))
})
```

Notice how the 3 above tests elegantly fail. The provided context ("validate width works") immediately indicates the culprit, which is a game changer for debugging.


## Testing JavaScript
We could not finish this chapter without mentionning tools to validate JavaScript code. We will particularly focus on [JSHint](https://jshint.com), which has a R [wrapper](https://github.com/dreamRs/jstools), nicely provided by the dreamRs team. Below is an example of how to check all the [shinyMobile](https://github.com/RinteRface/shinyMobile) JavaScript input bindings at once:

```{r}
bindings <- list.files(
  path = system.file("framework7-5.5.0/input-bindings", package = "shinyMobile"),
  recursive = TRUE,
  full.names = TRUE
)
jshint_file(input = bindings, options = jshint_options(jquery = TRUE, globals = list("Shiny", "app")))
```

You may fine tune the `jshint_file` behavior with the [`jshint_options`](https://jshint.com/docs/options/). One is often tempted to call `eval` in JS code, which will result in a JSHint error. An option called `evil` exists to disable the corresponding test. However, we recommand to play the game, accept those error and try to fix them instead of cheating with options!