# Adding more interactivity {#custom-templates-interactivity}
In this part, we are going to add more life to the template element. We first see how
to enhance an existing static HTML component. Then we will explore complex feedback mechanisms to provide more interactivity to your app.

## Custom progress bars
Progress bars are a good way to display metric related to a progress, for instance tracking the number of remaining tasks for a project. In general, those 
elements are static HTML. Hence, it would be interesting to be able to update the current
value from the server side. Since it is not an input, implementing an input binding is inaprioprioate, so we decide to proceed with the `sendCustomMessage`/`addCustomMessageHandler` pair. We first create the `tabler_progress`
tag which is mainly composed of:

  - `style` gives the current value of the progress. This is the main element
  - `min` and `max` are bounds, in general between 0 and 100
  - `id` ensure that our progress bar is unique so that we may update it without
  conflict

```{r}
tabler_progress <- function(id = NULL, value) {
  div(
    class = "progress",
    div(
      id = id, 
      class = "progress-bar",
      style = paste0("width: ", value, "%"),
      role = "progressbar",
      `aria-valuenow` = as.character(value),
      `aria-valuemin` = "0",
      `aria-valuemax` = "100",
      span(class = "sr-only", "38% Complete")
    )
  )
}

update_tabler_progress <- function(id, value, session = shiny::getDefaultReactiveDomain()) {
  message <- list(id = session$ns(id), value = value)
  session$sendCustomMessage(type = "update-progress", message)
}
```

The next element is the `update_tabler_progress` function which send 2 elements from R to JS:

  - The progress `id`
  - The new value

On the JS side, we have a basic `addCustomMessageHandler`. As mentionned in Chapter \@ref(survival-kit-javascript), `sendCustomMessage` and `addCustomMessageHandler` are connected by the same `type`. This is crucial! Moreover, as the sent message is a R list, 
it becomes an JSON, meaning that elements must be accessed with a `.`:

```javascript
$(function () {
  Shiny.addCustomMessageHandler('update-progress', function(message) {
    $('#' + message.id).css('width', message.value +'%');
  });
});
```

We finally test these components in a simple app:

```{r, eval=FALSE}
ui <- tabler_page(
  tags$head(
    tags$script(
      "$(function() {
        Shiny.addCustomMessageHandler('update-progress', function(message) {
          $('#' + message.id).css('width', message.value +'%');
        });
      });
      "
    )
  ),
  tabler_body(
    sliderInput(
      "progress_value", 
      "Progress value:",
      min = 0, 
      max = 100, 
      value = 50
    ),
    tabler_progress(id = "progress1", 12)
  )
)

server <- function(input, output, session) {
  observeEvent(input$progress_value, {
    update_tabler_progress(
      id = "progress1", 
      input$progress_value
    )
  })
}
shinyApp(ui, server)
```


<br>
<div class="callout callout-info">
NOTE: How to handle custom messages in shiny modules? Well, it is pretty straightforward: we wrap any id with the module namespace given by `session$ns()` before sending it to JS. You may even do it by default (without modules) like I did in the previous example since `session$ns()` will be `""`.
</div>
<br>


## Tab events
Do you remember about the navbar element and the tabsetpanel (also called navs) system of Chapter \@ref(custom-templates-skeleton)? Navs allow to organize any app into several tabs, acting like pages. This is a powerful tool for Shiny since it is currently not straightforward to create [multi-pages](https://community.rstudio.com/t/shiny-app-composed-of-many-many-pages/7698) Shiny apps like anyone would do with a website. Navs relies on the Bootstrap4 API but we only used few JS functions. There exist [events](https://getbootstrap.com/docs/4.0/components/navs/#events) that trigger after/before showing/hiding a tabs, that is `hidden.bs.tab` and `shown.bs.tab` (`hide.bs.tab` and `show.bs.tab` are triggers before). For instance, we may recover
the previously selected tab, namely `e.relatedTarget`, and store it in a shiny input with `Shiny.setInputValue`:

```javascript
$(function() {
  $('a[data-toggle="tab"]').on('shown.bs.tab', function (e) {
    Shiny.setInputValue('previous_tab', e.relatedTarget)
  });
});
```

