# Package creation {#custom-templates-package}

## Introduction
Translating an HTML template into an R API requires the creation of a package. As stated in Chapter \@ref(intro), this is not a good practice to proceed as follows:

```{r, eval=FALSE}
ui <- fluidPage(
  useShinydashboard(),
  tags$script(
    "$(function() {
      let boxBinding = new Shiny.InputBinding();
      $.extend(boxBinding, {
        initialize: function(el) {
          $(el).activateBox();
        },
        find: function(scope) {
          return $(scope).find('.box');
        },
        getValue: function(el) {
          let isCollapsed = $(el).hasClass('collapsed-box');
          return {collapsed: isCollapsed}; // this will be a list in R
        },
        setValue: function(el, value) {
          $(el).toggleBox();
        }, 
        receiveMessage: function(el, data) {
          this.setValue(el, data);
          $(el).trigger('change');
        },
        subscribe: function(el, callback) {
          $(el).on('click', '[data-widget=\"collapse\"]', function(event) {
            setTimeout(function() {
              callback();
            }, 550);
          }); 
          $(el).on('change', function(event) {
            setTimeout(function() {
              callback();
            }, 550);
          });
        },
        unsubscribe: function(el) {
          $(el).off('.boxBinding');
        }
      });
      Shiny.inputBindings.register(boxBinding, 'box-input');
    });
    "
  ),
  
  box2(
    title = textOutput("box_state"),
    "Box body",
    inputId = "mybox",
    collapsible = TRUE,
    plotOutput("plot")
  ),
  actionButton("toggle_box", "Toggle Box")
)

server <- function(input, output, session) {
  output$plot <- renderPlot({
      req(!input$mybox$collapsed)
      plot(rnorm(200))
    })
    
    output$box_state <- renderText({
      state <- if (input$mybox$collapsed) "collapsed" else "uncollapsed"
      paste("My box is", state)
    })
    
    observeEvent(input$toggle_box, {
      updateBox2("mybox")
    })
}

shinyApp(ui, server)
```

Imagine if we had to repeat the process for more than twenty components. The package structure provides many advantages like:

  - Develop a comprehensive documentation of the underlying API
  - Design unit tests to guaranty code robustness and improve long term plan
  - Relevant file organization
  
## Step by step
[{usethis}](https://usethis.r-lib.org) is a game changer for package creation.
Below is a quick overview of all steps I followed to setup this book side package.


```{r, eval = FALSE}
library(usethis)
create_package("mypk")
# TO FINISH
```