# Testing and validating templates elements {#custom-templates-testing}
Until now, we have been building the template boilerplates, that is the main skeleton functions (page, navbar, navbar menu, ...) as well as some components such as cards, ribbons, progress bars. We also exposed some techniques to substantially give more interactivity to the template, leveraging our freshly adquired JavaScript skills. Does this mean we are ready to make the template public? Not yet since some essentials steps are missing:


  - Input validation is a crucial step toward success. Briefly, it consists in checking user inputs so that your functions fail safely and elegantly by providing meaningful error messages or warnings. This concept has already been covered in R for Data Science, Advanced R and a lot of other resources. Hence, I am not trying to reinvent the wheel and we will rely on already existing patterns, whose effectiveness is no longer to be demonstrated. Welcome to the defensive programming world! 
  - On the other hand, testing components allows to check if a functions does what it is supposed to do. Consequently, it is a proof of robustness, and increases reproducibility. It significantly reduces the mental load when it comes to start code refactoring, thereby making you feel slightly less guilty about creating breaking changes since, most of the time, your tests will be able to capture those error. 



## Validate template functions
Below, we will show some examples to validate user inputs. We first consider the `tabler_card` element from Chapter \@ref(custom-templates-skeleton).

```{r}
tabler_card <- function(..., title = NULL, status = NULL, width = 6, stacked = FALSE, padding = NULL) {
  
  card_cl <- paste0(
    "card", 
    if (stacked) " card-stacked",
    if (!is.null(padding)) paste0(" card-", padding)
  )
  
  div(
    class = paste0("col-md-", width),
    div(
      class = card_cl,
      if (!is.null(status)) {
        div(class = paste0("card-status-top bg-", status))
      },
      div(
        class = "card-body",
        # we could have a smaller title like h4 or h5...
        if (!is.null(title)) {
          h3(class = "card-title", title)
        },
        ...
      )
    )
  )
  
}
```

The first thing is to think about what to validate. Here, I see at least 3 test:

  - checking the status value
  - checking the width value
  - checking the padding value
  
Writing validation tests requires knowledge about the underlying mechanisms. In our case, we must know the Bootstrap 4 grid rules and valid color statuses. The Bootstrap 4 [grid](https://getbootstrap.com/docs/4.0/layout/grid/) ranges from 1 to 12. In other words, a card having a width of 12 will take the full page wherease, we may align 3 cards or width 4. Valid [statuses](https://getbootstrap.com/docs/4.0/utilities/colors/) are primary, secondary, success, info, danger, warning, light and dark.

It is therefore pretty straightforward to validate the card width: it must be numeric and between 1 and 12. Moreover, since the template has other containers including the width parameter, we will create a function

```{r}
validate_width <- function(width) {
  if (is.numeric(width)) {
    if (width < 1 || width > 12) {
      stop("width must belong to [1, 12], as per Bootstrap 4 grid documentation. See more at https://getbootstrap.com/docs/4.0/layout/grid/")
    }
  } else {
    stop("width must be numeric")
  }
}

# validate_width(-1)
# validate_width(13)
# validate_width("string")
```

## Testing templates elements
