# htmltools overview {#htmltools-overview}

## HTML Tags
htmltools contain tags. However, by experience, htmltools contains more exported tags than shiny.
For instance, the HTML `<nav></nav>` tag, namely `tags$nav()` in R is not included in the shiny package but 
in htmltools. 

Within your package code, your tags will be like:

```{r, eval = FALSE}
# we use htmltools tags instead of shiny
htmltools::tags$div(...)
```

If you had to gather multiple tags together, prefer `tagList()` as `list()`, although the HTML output is the same. The first
has the shiny.tag.list class in addition to list.

## Notations
Whether to use `tags$div` or `div` is the tag is exported by default.
For instance, you could use `htmltools::div` but not `htmltools::nav` since nav does not 
have a dedicated function (only for p, h1, h2, h3, h4, h5, h6, a, br, div, span, pre, code, img, strong, em, hr). 
Rather use `htmltools::tags$nav`. Alternatively, there exists a function (in shiny and htmltools) 
called `withTags()`. Wrapping your code in this function enables you to use `withTags(nav(), ...)` 
instead of `tags$nav()`.


## Alternative way to write tags
htmltools and shiny come with the `HTML()` function that you can feed with raw HTML:

```{r, eval = FALSE}
HTML('<div>Blabla</div>')
# will render exactly like
div("Blabla")

# but there class is different
class(HTML('<div>Blabla</div>'))
class(div("Blabla"))
```

You will not be able to use tag related functions, as in the following parts.
Therefore, I strongly recommand using R and not mixing HTML in R.

## Playing with tags

### Tags structure

According to the `tag` function, a tag has:
- a name such as span, div, h1 ... `tag$name`
- some attributes, which you can access with `tag$attribs`
- children, which you can access with `tag$children`
- a class, namely "shiny.tag"

For instance:

```{r, eval = FALSE}
# create the tag
myTag <- div(
  class = "divclass", 
  id = "first",
  h1("Here comes your baby"),
  span(class = "child", id = "baby", "Crying")
)
# access its name
myTag$name
# access its attributes (id and class)
myTag$attribs
# access children (returns a list of 2 elements)
myTag$children
# access its class
class(myTag)
```

How to modify the class of the second child, namely span?

```{r, eval = FALSE}
second_children <- myTag$children[[2]]
second_children$attribs$class <- "adult"
myTag
# Hummm, this is not working ...
```

Why is this not working? By assigning `myTag$children[[2]]` to second_children, `second_children$attribs$class <- "adult"` modifies the class of the copy and not the original object. Thus we do:

```{r, eval = FALSE}
myTag$children[[2]]$attribs$class <- "adult"
myTag
```

In the following section we explore helper functions, such as `tagAppendChild` from htmltools.


### Useful functions for tags

htmltools and Shiny have powerful functions to easily add attributes to tags, check for existing attributes, get attributes and add other siblings to a list of tags.

#### Add attributes

- `tagAppendAttributes`: this function allow you to add a new attribute to the current tag.

For instance, assuming you created a div for which you forgot to add an id attribute:

```{r, eval = FALSE}
mydiv <- div("Where is my brain")
mydiv <- tagAppendAttributes(mydiv, id = "here_it_is")
```

You can pass as many attributes as you want, including non standard attributes such as `data-toggle` (see Bootstrap 3 tabs for instance):

```{r, eval = FALSE}
mydiv <- tagAppendAttributes(mydiv, `data-toggle` = "tabs")
# even though you could proceed as follows
mydiv$attribs[["aria-controls"]] <- "home"
```

#### Check if tag has specific attribute

- `tagHasAttribute`: to check if a tag has a specific attribute

```{r, eval = FALSE}
# I want to know if div has a class
mydiv <- div(class = "myclass")
has_class <- tagHasAttribute(mydiv, "class")
has_class
# if you are familiar with %>%
has_class <- mydiv %>% tagHasAttribute("class")
has_class
```

#### Get all attributes 
 
- `tagGetAttribute`: to get the value of the targeted attributes, if it exists, otherwise NULL.

```{r, eval = FALSE}
mydiv <- div(class = "test")
# returns the class
tagGetAttribute(mydiv, "class")
# returns NULL
tagGetAttribute(mydiv, "id")
```

#### Set child/children

- `tagSetChildren` allows to create children for a given tag. For instance:

```{r, eval = FALSE}
mydiv <- div(class = "parent", id = "mother", "Not the mama!!!")
# mydiv has 1 child "Not the mama!!!"
mydiv 
children <- lapply(1:3, span)
mydiv <- tagSetChildren(mydiv, children)
# mydiv has 3 children, the first one was removed
mydiv 
```

Notice that `tagSetChildren` removes all existing children. Below we see another set of functions to add children while conserving existing ones.

#### Add child or children

- `tagAppendChild` and `tagAppendChildren`: add other tags to an existing tag.
Whereas `tagAppendChild` only takes one tag, you can pass a list of tags to `tagAppendChildren`.

```{r, eval = FALSE}
mydiv <- div(class = "parent", id = "mother", "Not the mama!!!")
otherTag <- span("I am your child")
mydiv <- tagAppendChild(mydiv, otherTag)
```

You might wonder why there is no `tagRemoveChild` or `tagRemoveAttributes`.
Let's look at the `tagAppendChild`

```{r, eval = FALSE}
tagAppendChild <- function (tag, child) {
  tag$children[[length(tag$children) + 1]] <- child
  tag
}
```

Below we write the `tagRemoveChild`, where tag is the target and n is the position to remove in the list of children:

```{r, eval = FALSE}
mydiv <- div(class = "parent", id = "mother", "Not the mama!!!", span("Hey!"))

# we create the tagRemoveChild function
tagRemoveChild <- function(tag, n) {
  # check if the list is empty
  if (rlang::is_empty(tag$children)) {
    stop(paste(tag$name, "does not have any children"))
  }
  tag$children[n] <- NULL
  tag
}
mydiv <- tagRemoveChild(mydiv, 1)
mydiv
```

When defining the `tagRemoveChild`, we choose `[` instead of `[[` to allow to select multiple list elements:

```{r, eval = FALSE}
mydiv <- div(class = "parent", id = "mother", "Not the mama!!!", "Hey!")
# fails
`[[`(mydiv$children, c(1, 2))
# works
`[`(mydiv$children, c(1, 2))
```

Alternatively, we could also create a `tagRemoveChildren` function. Also notice that the function raises an error if the provided tag does not have children. 

### Other interesting functions
The [brighter](https://github.com/ThinkR-open/brighter) package written by Colin Fay contains neat functions to edit your tags. Particularly, the `tagRemoveAttributes`

```{r, eval = FALSE}
remotes::install_github("Thinkr-open/brighter")
library(brighter)
```

```{r, eval = FALSE}
mydiv <- div(class = "test", id = "coucou", "Hello")
tagRemoveAttributes(mydiv, "class", "id")
