# Dependency utilities {#htmltools-dependencies}
When creating a new template, you sometimes need to import custom HTML dependencies
that do not come along with shiny. No problem, htmltools is here for you (shiny also 
contains these functions).

```{r setup}
library(shiny)
library(shinydashboard)
```

## The dirty approach
Let's consider the following example. I want to include a bootstrap 4 card in a shiny app.
This example is taken from an interesting question [here](https://community.rstudio.com/t/create-a-div-using-htmltools-withtags/22439/2).
The naive approach would be to include the HTML code directly in the app code

```{r, eval = FALSE}

# we create the card function before
my_card <- function(...) {
  htmltools::withTags(
    div(
      class = "card border-success mb-3",
      div(class = "card-header bg-transparent border-success"),
      div(
        class = "card-body text-success",
        h3(class = "card-title", "title"),
        p(class = "card-text", ...)
      ),
      div(class = "card-footer bg-transparent border-success", "footer")
    )
  )
}

# we build our app
shinyApp(
  ui = fluidPage(
    fluidRow(
      column(
        width = 6,
        align = "center",
        br(),
        my_card("blablabla. PouetPouet Pouet.")
      )
    )
  ),
  server = function(input, output) {}
)
```

and desesperately see that nothing is displayed. If you remember, this was expected since
shiny does not contain bootstrap 4 dependencies and this card is unfortunately a
bootstrap 4 object. Don't panic! We just need to tell shiny to load the css we need to display
this card (if required, we could include the javascript as well). We could use either
`includeCSS()`, `tags$head(tags$link(rel = "stylesheet", type = "text/css", href = "custom.css"))`. See
more [here](https://shiny.rstudio.com/articles/css.html).

```{r, eval = FALSE}
shinyApp(
  ui = fluidPage(
    # load the css code
    includeCSS(path = "https://maxcdn.bootstrapcdn.com/bootstrap/4.0.0/css/bootstrap.min.css"),
    fluidRow(
      column(
        width = 6,
        align = "center",
        br(),
        my_card("blablabla. PouetPouet Pouet.")
      )
    )
  ),
  server = function(input, output) {}
)
```

The card is ugly (which is another problem we will fix later) but at least displayed.

When I say this approach is dirty, it is because it will not be easily re-usable by others.
Instead, we prefer a packaging approach, like in the next section.

## The clean approach

We will use the `htmlDependency` and `attachDependencies` functions from htmltools.
The htmlDependency takes several arguments:

- the name of your dependency
- the version (useful to remember on which version it is built upon)
- a path to the dependency (can be a CDN or a local folder)
- script and stylesheet to respectively pass css and scripts

```{r, eval = FALSE}
# handle dependency
card_css <- "bootstrap.min.css"
bs4_card_dep <- function() {
  htmltools::htmlDependency(
    name = "bs4_card",
    version = "1.0",
    src = c(href = "https://maxcdn.bootstrapcdn.com/bootstrap/4.0.0/css/"),
    stylesheet = card_css
  )
}
```

We create the card tag and give it the bootstrap 4 dependency through the `attachDependencies()` function. In recent version of htmltools, we may simply use
`tagList(tag, deps)` instead.

```{r, eval = FALSE}
# create the card
my_card <- function(...) {
  cardTag <- htmltools::withTags(
    div(
      class = "card border-success mb-3",
      div(class = "card-header bg-transparent border-success"),
      div(
        class = "card-body text-success",
        h3(class = "card-title", "title"),
        p(class = "card-text", ...)
      ),
      div(class = "card-footer bg-transparent border-success", "footer")
    )
  )
  
  # attach dependencies (old way)
  # htmltools::attachDependencies(cardTag, bs4_card_dep())
  
  # simpler way
  tagList(cardTag, bs4_card_dep())
  
}

```

We finally run our app:

```{r, eval = FALSE}
# run shiny app 
ui <- fluidPage(
  title = "Hello Shiny!",
  fluidRow(
    column(
      width = 6,
      align = "center",
      br(),
      my_card("blablabla. PouetPouet Pouet.")
    )
  )
)

shinyApp(ui, server = function(input, output) { })
```

With this approach, you could develop a package of custom dependencies that people
could use when they need to add custom elements in shiny.


## Another example: Importing HTML dependencies from other packages

You may know shinydashboard, a package to design dashboards with shiny. In the following, we would like to integrate the box component in a classic Shiny App (without the dashboard layout). However, if you try to include the Shinydashboard box tag, you will notice that nothing is displayed since Shiny does not have shinydashboard dependencies. Fortunately htmltools contains a function, namely `findDependencies` that looks for all dependencies attached to a tag. How about extracting shinydashboard dependencies? Before going futher, let's define the basic skeleton of a shinydashboard:

```{r, eval = FALSE}
shinyApp(
  ui = dashboardPage(
    dashboardHeader(),
    dashboardSidebar(),
    dashboardBody(),
    title = "Dashboard example"
  ),
  server = function(input, output) { }
)
```

We don't need to understand shinydashboard details. However, if you are interested to dig in, [help yourself](https://rstudio.github.io/shinydashboard/). What is important here is the main
wrapper function `dashboardPage`. (You should already be familiar with `fluidPage`, another wrapper function). We apply `findDependencies` on `dashboardPage`.

```{r, eval = FALSE}
deps <- findDependencies(
  shinydashboard::dashboardPage(
    header = shinydashboard::dashboardHeader(), 
    sidebar = shinydashboard::dashboardSidebar(), 
    body = shinydashboard::dashboardBody()
  )
)
deps
```

deps is a list containg 4 dependencies:

- [Font Awesome](https://fontawesome.com) handles icons
- [Bootstrap](https://getbootstrap.com/docs/3.3/) is the main HTML/CSS/JS template. Importantly,
please note the version 3.3.7, whereas the current is 4.3.1
- [AdminLTE](https://adminlte.io) is the dependency containg HTML/CSS/JS related to the admin template.
It is closely linked to Bootstrap 3. 
- shinydashboard, the CSS and javascript necessary for shinydashboard to work properly. In practice,
integrating custom HTML templates to shiny does not usually work out of the box for many reasons (Explain why!) and some modifications are necessary.


```
[[1]]
List of 10
$ name      : chr "font-awesome"
$ version   : chr "5.3.1"
$ src       :List of 1
..$ file: chr "www/shared/fontawesome"
$ meta      : NULL
$ script    : NULL
$ stylesheet: chr [1:2] "css/all.min.css" "css/v4-shims.min.css"
$ head      : NULL
$ attachment: NULL
$ package   : chr "shiny"
$ all_files : logi TRUE
- attr(*, "class")= chr "html_dependency"
[[2]]
List of 10
$ name      : chr "bootstrap"
$ version   : chr "3.3.7"
$ src       :List of 2
..$ href: chr "shared/bootstrap"
..$ file: chr "/Library/Frameworks/R.framework/Versions/3.5/Resources/library/shiny/www/shared/bootstrap"
$ meta      :List of 1
..$ viewport: chr "width=device-width, initial-scale=1"
$ script    : chr [1:3] "js/bootstrap.min.js" "shim/html5shiv.min.js" "shim/respond.min.js"
$ stylesheet: chr "css/bootstrap.min.css"
$ head      : NULL
$ attachment: NULL
$ package   : NULL
$ all_files : logi TRUE
- attr(*, "class")= chr "html_dependency"
[[3]]
List of 10
$ name      : chr "AdminLTE"
$ version   : chr "2.0.6"
$ src       :List of 1
..$ file: chr "/Library/Frameworks/R.framework/Versions/3.5/Resources/library/shinydashboard/AdminLTE"
$ meta      : NULL
$ script    : chr "app.min.js"
$ stylesheet: chr [1:2] "AdminLTE.min.css" "_all-skins.min.css"
$ head      : NULL
$ attachment: NULL
$ package   : NULL
$ all_files : logi TRUE
- attr(*, "class")= chr "html_dependency"
[[4]]
List of 10
$ name      : chr "shinydashboard"
$ version   : chr "0.7.1"
$ src       :List of 1
..$ file: chr "/Library/Frameworks/R.framework/Versions/3.5/Resources/library/shinydashboard"
$ meta      : NULL
$ script    : chr "shinydashboard.min.js"
$ stylesheet: chr "shinydashboard.css"
$ head      : NULL
$ attachment: NULL
$ package   : NULL
$ all_files : logi TRUE
- attr(*, "class")= chr "html_dependency"
```

Below, we attach the dependencies to the `box` with `tagList`, as shown above. Notice that our custom `box` does not contain all parameters from shinydashboard but this is not what matters in this example.

```{r, eval = FALSE}
my_box <- function(title, status) {
  tagList(box(title = title, status = status), deps)
}
ui <- fluidPage(
  titlePanel("Shiny with a box"),
  my_box(title = "My box", status = "danger"),
)
server <- function(input, output) {}
shinyApp(ui, server)
```

Now, you may imagine the possibilities are almost unlimited! 