--- 
title: "Outstanding User Interfaces with Shiny"
author: "David Granjon"
date: "`r Sys.Date()`"
documentclass: krantz
bibliography: [book.bib, packages.bib, externalPackages.bib]
biblio-style: apalike
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
site: bookdown::bookdown_site
description: "A book about deeply customizing Shiny app for production."
favicon: images/favicon.ico
github-repo: DivadNojnarg/outstanding-shiny-ui
graphics: yes
nocite: '@*'
---

# Welcome {-}

This book is still under active development and intended for a Q2 release in the [R Series](https://www.routledge.com/go/the-r-series) by Chapman
& Hall.

# Prerequisites {-}

Before going further, the reader must:

  - Be comfortable with R structures, functional and object oriented programming ([R6](https://r6.r-lib.org/))
  - Have solid knowledge of [Shiny](https://mastering-shiny.org)
  
  
Basic knowledge in HTML and JavaScript is a plus but not mandatory.

## Disclaimer {-}

This book is not an HTML/Javascript/CSS course and won't make you become a web developer! Instead, it provides a _survival kit_ to customize Shiny. I am sure however that readers will want to explore more about these topics.


## Is this book for me? {-}

If you fulfill the mentioned prerequisites, you should read this book if you answer yes to the following questions:

  - Do you want to know how to develop shiny apps with a professional look and feel?
  - Have you ever wondered how to design new input widgets?
  - Are you simply curious about what happens under the hood?
  

## Learning objectives

This book will help you to:

  - Manipulate R Shiny tags with to create custom layouts
  - Understand how web dependencies are handled
  - Discover the steps to create a new template from scratch
  - Learn how shiny deals with inputs
  - Learn more about less documented Shiny mechanisms (websockets, sessions, ...)

## Related content {-}

[Novartis](https://www.novartis.com) associates Mustapha Larbaoui and David Granjon gave a workshop entitled "Advanced User Interfaces for Shiny Developers" at the latest 2020 virtual [e-Rum](https://2020.erum.io/program/workshops/) conference whose detailed agenda is available [here](https://github.com/Novartis/Advanced-User-Interfaces-for-Shiny-Developers). This two hours workshop takes a lot of inspiration on this book. An [RStudio Cloud](https://rstudio.cloud) dedicated project is deployed [here](https://rstudio.cloud/project/1395473).

## Acknowledgements {-}
- I am very grateful to Douglas Robinson for proof reading the book and fixing many typos it contained
- A special thanks to my friends John and Victor for contributing to [RinteRface](https://rinterface.com/)
- Thanks to the eRum organizers for inviting me to the virtual [e-Rum2020](https://2020.erum.io) conference
- RinteRface and this book won't exist without the amazing R community. Thanks for their valuable feedback


## Packages {-}
```{r setup}
library(shiny)
library(shinydashboard)
library(cascadess)
library(htmltools)
library(purrr)
library(magrittr)
library(ggplot2)
library(thematic)
library(fresh)
library(testthat)
library(jstools)
library(scales)
library(dplyr)
library(apexcharter)
library(shinyWidgets)
```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
