--- 
title: "Outstanding User Interfaces with Shiny"
author: "David Granjon"
date: "`r Sys.Date()`"
documentclass: krantz
bibliography: [book.bib, packages.bib, externalPackages.bib]
biblio-style: apalike
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
site: bookdown::bookdown_site
description: "A book about deeply customizing Shiny app for production."
favicon: images/favicon.ico
github-repo: DivadNojnarg/outstanding-shiny-ui
graphics: yes
nocite: '@*'
---

# Prerequisites {-}

  - Be familiar with [Shiny](https://mastering-shiny.org)
  - Basic knowledge in HTML and JavaScript is a plus but not mandatory

## Disclaimer {-}

This book is not an HTML/Javascript/CSS course! Instead, it provides a _survival kit_ to be able to customize Shiny. I am sure however that readers will want to explore more about these topics.

## Is this book for me? {-}

You should read this book if you answer yes to the following questions:

  - Do you want to know how to develop outstanding shiny apps?
  - Have you ever wondered how to develop new input widgets?
  

## Related content {-}

See the [RStudio Cloud](https://rstudio.cloud) dedicated project.
This book will serve as a reference for the e-Rum2020 [workshop](https://2020.erum.io/program/workshops/) about "Advanced User Interfaces for Shiny Developers", provided by [Novartis](https://www.novartis.com). The detailed agenda is available [here](https://github.com/Novartis/Advanced-User-Interfaces-for-Shiny-Developers).

## Acknowledgements {-}
- I am very grateful to Douglas Robinson for proof reading the book and fixing many typos it contained.
- A special thanks to my friends John and Victor for contributing to RinteRface. By contribution, I am not meaning only code contribution but also support in any form.
- Thanks to the eRum organizers for giving me the opportunity to present this work at the virtual [e-Rum2020](https://2020.erum.io) conference
- RinteRface and this book won't exist without the amazing R community. Thanks for their valuable feedback.


## Packages {-}
```{r setup}
library(shiny)
library(shinydashboard)
library(cascadess)
library(htmltools)
library(purrr)
library(magrittr)
library(ggplot2)
library(thematic)
library(fresh)
library(testthat)
library(jstools)
library(scales)
library(dplyr)
library(apexcharter)
library(shinyWidgets)
```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
