# Design widgets {#mobile-widgets}
Framework7 brings dozen of different widgets like a photo browser, virtual lists (high performance lists),
messages, notifications, toasts, ...

Looking at the [documentation](https://v5.framework7.io/docs/), the API is most of the time
always the same that is, we create the widget:

```javascript
app.widget.create(parameters);
```

and we update, open or close it later:

```javascript
app.widget.update(newParameters);
app.widget.open();
app.widget.close();
```

I must admit, there are few deviations like the navbar (`app.navbar.show()`) or the dialog but we have enough common points to
design a main wrapper that creates any widget and update/open/close it.

What we do below significantly simplify the R/JS API at a price to slightly reduce the
customization. 

## Build the UI
We know that JavaScript must receive a configuration object to create the widget instance.
As shown earlier in this book, there is a simple way to achieve this. Let's consider the [gauge](https://v5.framework7.io/docs/gauge.html) example. 

On the UI side, we expect to have:

```html
<div class="gauge"></div>
```

Upon instantiating, Framework7 populates this container with the relevant tags.
The `f7_gauge` function creates a `div` tag with the `gauge` class as well as a configuration
tag:

```{r, eval=FALSE}
f7_gauge <- function(id, value, options = NULL) {

  if (is.null(options[["valueText"]])) options[["valueText"]] <- paste(value * 100, "%")

  gaugeProps <- c(list(value = value), options)

  gaugeConfig <- shiny::tags$script(
    type = "application/json",
    `data-for` = id,
    jsonlite::toJSON(
      x = gaugeProps,
      auto_unbox = TRUE,
      json_verbatim = TRUE
    )
  )

  shiny::tags$div(
    class = "gauge",
    id = id,
    gaugeConfig
  )
}
```

We provide a default for the `valueText` option that should display the current
value followed by a `%` symbol. Note that the class is crucial to target the relevant tag on the JS side. 
All other widgets will moreless follow the same scheme. Be careful about partial matching
with the `$` operator. This is the reason why we used `[[` instead: with `$`, `valueText` would be
matched with `valueTextColor`, leading to misleading behavior.


## Widgets without preexisting UI
There are few widgets like toasts, notifications that don't have any predefined UI
element when the app starts. In this case, we simply send the configuration to JS, through
the `session`:

```{r, eval=FALSE}
f7_notif <- function(id = NULL, text, options = NULL, session = shiny::getDefaultReactiveDomain()) {

  if (!is.null(options$icon)) options$icon <- as.character(options$icon)

  message <- c(dropNulls(list(id = id, text = text)), options)
  # see my-app.js function
  sendCustomMessage("notification", message, session)

}
```

Pay attention to the `icon` element: we can't convert shiny tags to JSON, so they
must be converted to character first. If multiple parameters should contain tags,
you must treat them accordingly!

## Initialize the widget
On the JS side, we create a new script, `widgets.js` 

```{r, eval=FALSE}
library(charpente)
create_js("widgets")
```

We set an array containing all compatible widget names in two categories and concatenate
in a `widgets` element:

```javascript
const uiWidgets = ["gauge", "swiper", "searchbar"];
const serverWidgets = ["toast", "photoBrowser", "notification"];
const widgets = uiWidgets.concat(serverWidgets);
```

We then define the `activateWidget` function, only considering UI widgets. 
Since we have two widgets categories, this function first checks whether the widget is part of the `uiWidgets` array:

```javascript
if (uiWidgets.indexOf(widget) > -1) {
  // Do things
}
```

As there may be multiple widgets of the same time to initialize, we must loop through all
possible elements. This is where the class is important and must match the widget name.
For instance, the gauge has the `gauge` class and the methods are always `app.gauge.`.
How do we loop through multiple widgets? We use the jQuery `each` method:

```javascript
if (uiWidgets.indexOf(widget) > -1) {
  $("." + widget).each(function() {
    // Do things
  }
}
```

We see that `$("." + widget)` gives `$(".gauge)` when the widget is a gauge, which
will target all gauges one by one. Then for each gauge, we extract the configuration containing
all options passed by the end user. Remember that each element has a unique id. 
We extract the current element `$(this)` in the `$el` variable and search for a 
script tag pointing to the unique tag having `$el.attr("id")` as id. The configuration is parsed
to convert it to an object. Note that most of the time, Framework7 expects to have a `el`
attributes which simply contains the CSS selector of the current element, 
in other words its unique id `'#' + $el.attr("id")`:

```javascript
if (uiWidgets.indexOf(widget) > -1) {
  $("." + widget).each(function() {
    var $el = $(this);
    var config = $(document).find(
      "script[data-for='" + $el.attr("id") + "']"
    );
    config = JSON.parse(config.html());
    // add the id
    config.el = '#' + $el.attr("id");
  }
}
```

The final step is to initialize the widget, which is quite straightforward if we notice that
`app.gauge` is the same as `app["gauge"]`. We obtain the general code:

```javascript
app[widget].create(config);
```

For the server widgets, it is even simpler. We recover the message with a `Shiny.addCustomMessageHandler("type", callback)` and
initialize it. The only possible source of problem is the custom message `type` that must be the same 
as the one specified in the R function with `session$sendCustomMessage("type", message)`. We create
an `else` statement following the `if` condition and put that code inside:

```javascript
Shiny.addCustomMessageHandler(widget, function(message) {
  if (message.id !== undefined) {
    message.on = {
      opened: function() {
        Shiny.setInputValue(message.id, true);
      },
      closed: function() {
        Shiny.setInputValue(message.id, false);
        app.data[widget][message.id].destroy();
      }
    }; 
  }

  app[widget].create(message).open();
});
```

As shown in the above code, we can also immediately open the widget by chaining methods. Moreover, it 
is always good practice to let Shiny know about the widget state, that is whether it is currently
opened. This is the reason why we added a `on` property to the message. All widgets trigger events, 
for instance notifications have the `notification:opened` and `notification:closed`. For each event, 
we set an input value on the fly, with `Shiny.setInputValue`. This way, the template users are
able to know exactly when the widget is closed or opened and trigger any subsequent action.
This obviously requires the widget to pass an optional `id` attribute to ensure the uniqueness!


The full JavaScript code may be found below.

```javascript
// Instantiate a widget
activateWidget = function(widget) {
  // Handle ui side widgets
  if (uiWidgets.indexOf(widget) > -1) {
    $("." + widget).each(function() {
      var $el = $(this);
      var config = $(document).find(
        "script[data-for='" + $el.attr("id") + "']"
      );
      config = JSON.parse(config.html());
      // add the id
      config.el = '#' + $el.attr("id");

      // feed the create method
      app[widget].create(config);
    });
  } else {
    // This concerns toasts, notifications, photoBrowser, ...
    // that don't have any UI element in the DOM before creating
    // the widget instance.
    Shiny.addCustomMessageHandler(widget, function(message) {
      if (message.id !== undefined) {
        message.on = {
          opened: function() {
            Shiny.setInputValue(message.id, true);
          },
          closed: function() {
            Shiny.setInputValue(message.id, false);
            app.data[widget][message.id].destroy();
          }
        }; 
      }
      app[widget].create(message).open();
    });
  }
};
```

The final step is to activate all widgets. We proceed with a loop:

```javascript
// Loop over all widgets to activate them
widgets.forEach(function(w) {
  activateWidget(w);
});
```

Let's try below with a notification example, where we capture the state of the notification
in an input element:

```{r, eval=FALSE}
library(shiny)
ui <- f7_page(
  navbar = f7_navbar("Title"),
  toolbar = f7_toolbar(),
  title = "shinyMobile",
  options = list(
    theme = "ios",
    version = "1.0.0",
    taphold = TRUE,
    color = "#42f5a1",
    filled = TRUE,
    dark = TRUE
  )
)

server <- function(input, output, session) {
  observe({
    f7_notif(id = "welcome", "Helloooooo", options = list(closeTimeout = 2000))
  })
  
  observeEvent(input$welcome, {
    shiny::showNotification(sprintf("Notification is %s", input$welcome))
  })
}

shinyApp(ui, server)
```

## Update widgets
We would like to develop a similar generalized interface to update any element in the DOM.
Instead of having `update_f7_gauge`, `update_f7_swiper`, ... we want an `update_f7_instance` function. 

We leverage the `app.data` element that stores all instances by widget type. In the previous part,
we already created a cache for tooltips, so let's do it for gauges:

```javascript
config.data = function() {
  return {
    // any other widget type to cache ...
    gauge: []
  };
};
```

At that step, the array name must be the same as the app method. For instance,
we have `app.gauge`, which means that we should create `config.data.gauge` and not `config.data.gauges`,
as it would lead to errors later.

Once the cache is available, we have to modify the JavaScript that creates the widget instance,
to store the new instance in the cache. We add the following code, where w refers to the widget instance:

```javascript
// ui widgets
app.data[widget][$el.attr("id")] = w;
```

This manipulation does not make sense for server widgets!

The `activateWidget` function should be:

```javascript
// Instantiate a widget
activateWidget = function(widget) {
  // Handle ui side widgets
  if (uiWidgets.indexOf(widget) > -1) {
    $("." + widget).each(function() {
      var $el = $(this);
      var config = $(document).find(
        "script[data-for='" + $el.attr("id") + "']"
      );
      config = JSON.parse(config.html());
      // add the id
      config.el = '#' + $el.attr("id");

      // feed the create method
      var w = app[widget].create(config);
      // Store the widget instance in the app data cache
      app.data[widget][$el.attr("id")] = w;
    });
  } else {
    // This concerns toasts, notifications, photoBrowser, ...
    // that don't have any UI element in the DOM before creating
    // the widget instance.
    Shiny.addCustomMessageHandler(widget, function(message) {
      if (message.id !== undefined) {
        message.on = {
          opened: function() {
            Shiny.setInputValue(message.id, true);
          },
          closed: function() {
            Shiny.setInputValue(message.id, false);
            app.data[widget][message.id].destroy();
          }
        }; 
      }
      var w = app[widget].create(message);
      w.open();
    });
  }
};
```

Once done, this is time to design `update_f7_instance`. The R code sends a message to
the current session containing:

  - The id of the element to update.
  - The new configuration.
  
Since we send a JSON, the hardest part is to correctly process shiny tags. How do we
track shiny tags. As a reminder, let's run the code below:

```{r}
class(shiny::div())
class(shiny::tagList(shiny::div(), shiny::h1()))
```

For each configuration element, we must check whether its class contains `shiny.tag` or `shiny.tag.list` and convert it to a character. Moreover, it may contain a nested list like this:

```{r, eval = FALSE}
options = list(
  buttons = list(
   list(
     text = "Some text",
     icon = f7Icon("info"),
     color = "pink"
   )
  )
)
```

In that case, our function must be recursive and call itself whenever an item having the `list` class
is found. If the element is simple text, numeric, we return it as is.

We finally get:

```{r, eval=FALSE}
update_f7_instance <- function(id, options, session = shiny::getDefaultReactiveDomain()) {

  # Convert any shiny tag into character so that toJSON does not cry
  listRenderTags <- function(l) {
    lapply(
      X = l,
      function(x) {
        if (inherits(x, c("shiny.tag", "shiny.tag.list"))) {
          as.character(x)
        } else if (inherits(x, "list")) {
          # Recursive part
          listRenderTags(x)
        } else {
          x
        }
      }
    )
  }
  options <- listRenderTags(options)

  message <- list(id = id, options = options)
  sendCustomMessage("update-instance", message, session)
}
```

On the JS side, we receive the message (still in the `widget.js` script):

```javascript
Shiny.addCustomMessageHandler('update-instance', function(message) {
  // Treat message ...
});
```

All widgets are stored by type in the app data, for instance, the element having for unique id
`mygauge` is located in `app.data["gauge"]["mygauge"]`. As there is no easy way to recover the widget type
given its id, the first step of the message handler is to find where our instance is located. We design a
nested for loop. The outer loop scans all app.data properties (ie widget categories) and the inner loop
scans all existing instances for each category. Whenever, the `message.id` matches the instance
name, we store the corresponding widget category in a variable:

```javascript
var instanceFamily;
for (const property in app.data) {
  for (const e in app.data[property]) {
    if (e === message.id) {
      instanceFamily = property;
    }
  }
}
```

We then access the old instance using the newly defined variable and the `message.id`.
We capture its parameters located in `oldInstance.params`. From there, multiple options are available:

  - We extend the old configuration by the new one.
  - We entirely overwrite the existing options. 
  
Below, we decided to merge the old and new configurations using `app.utils.extend`:

```javascript
var oldInstance = app.data[instanceFamily][message.id];
var oldConfig = oldInstance.params;
var newConfig = app.utils.extend(oldConfig,  message.options);
```

The next step is to destroy the old instance, initialize the new instance and refresh 
the `app.data` cache:

```javascript
// Destroy old instance
oldInstance.destroy();
// Create new config
var newInstance = app[instanceFamily].create(newConfig);
// Update app data
app.data[instanceFamily][message.id] = newInstance;
```

The whole code is located below:

```javascript
Shiny.addCustomMessageHandler('update-instance', function(message) {
  // Recover in which array is stored the given instance.
  // Uniqueness is ensured since HTML id are supposed to be unique.
  var instanceFamily;
  for (const property in app.data) {
    for (const e in app.data[property]) {
      if (e === message.id) {
        instanceFamily = property;
      }
    }
  }

  var oldInstance = app.data[instanceFamily][message.id];
  var oldConfig = oldInstance.params;
  var newConfig = app.utils.extend(oldConfig,  message.options);

  // Destroy old instance
  oldInstance.destroy();
  // Create new config
  var newInstance = app[instanceFamily].create(newConfig);
  // Update app data
  app.data[instanceFamily][message.id] = newInstance;
});
```

Below is an example showing how to update a gauge from the server. As you may notice, this
approach is not perfect as the user has to explicitly update the `valueText` field so that
it reflects the new value. Similarly, you may ask why the gauge value has to be between 0 and 1,
instead of 0 and 100. The reason comes from the Framework7 API. One might be tempted to convert the value inside `f7_gauge` (so that the user only provides number between 0 and 100), but
this would also mean to manually convert the value in the `update_f7_instance` function later. As stated in previous chapters, there is always a compromise to do between a simple API easy to maintain (for the developer) and user experience. This issue may/should be solved by a comprehensive documentation.

```{r, eval=FALSE}
ui <- f7_page(
  f7_gauge(
    "mygauge", 
    value = 0.1,
    options = list(
      type  = "semicircle",
      borderColor = "#2196f3",
      borderWidth = 10,
      valueFontSize = 41,
      valueTextColor = "#2196f3",
      labelText = "amount of something"
    )
  ),
  navbar = f7_navbar("Title"),
  toolbar = f7_toolbar(),
  title = "shinyMobile",
  options = list(
    theme = "ios",
    version = "1.0.0",
    taphold = TRUE,
    color = "#42f5a1",
    filled = TRUE,
    dark = TRUE
  )
)

server <- function(input, output, session) {
  observe({
    Sys.sleep(2)
    update_f7_instance(
      "mygauge", 
      options = list(
        value = 0.75, 
        valueText = "75 %", 
        labelText = "New label!"
      )
    )
  })
}

shinyApp(ui, server)
```
