# Design widgets {#mobile-widgets}
Framework7 brings dozen of different widgets like a photo browser, virtual lists (high performance lists),
messages, notifications, toasts, ...

Looking at the [documentation](https://v5.framework7.io/docs/), the API is most of the time
always the same that is, we create the widget:

```javascript
app.widget.create(parameters);
```

and we update, open or close it later:

```javascript
app.widget.update(newParameters);
app.widget.open();
app.widget.close();
```

I must admit, there are few deviations like the navbar (`app.navbar.show()`) or the dialog but we have enough common points to
design a main wrapper that creates any widget and update/open/close it.

What we do below significantly simplify the R/JS API at a price to slightly reduce the
customization. 

## Build the UI
We know that JavaScript must receive a configuration object to create the widget instance.
As shown earlier in this book, there is a simple way to achieve this. Let's consider the [gauge](https://v5.framework7.io/docs/gauge.html) example. 

On the UI side, we expect to have:

```html
<div class="gauge"></div>
```

Upon instantiating, Framework7 populates this container with the relevant tags.
The `f7_gauge` function creates a `div` tag with the `gauge` class as well as a configuration
tag:

```{r, eval=FALSE}
f7Gauge <- function(id, value, options = NULL) {

  if (is.null(options$valueText)) options$valueText <- paste(options$value, "%")

  gaugeProps <- c(list(value = value), options)

  gaugeConfig <- shiny::tags$script(
    type = "application/json",
    `data-for` = id,
    jsonlite::toJSON(
      x = gaugeProps,
      auto_unbox = TRUE,
      json_verbatim = TRUE
    )
  )

  shiny::tags$div(
    class = "gauge",
    id = id,
    gaugeConfig
  )
}
```

We provide a default for the `valueText` option that should display the current
value followed by a `%` symbol. Note that the class is crucial to target the relevant tag on the JS side. 
All other widgets will moreless follow the same scheme.


## Widgets without preexisting UI
There are few widgets like toasts, notifications that don't have any predefined UI
element when the app starts. In this case, we simply send the configuration to JS, through
the `session`:

```{r, eval=FALSE}
f7Notif <- function(text, options = NULL, session = shiny::getDefaultReactiveDomain()) {

  if (!is.null(options$icon)) options$icon <- as.character(options$icon)

  message <- c(list(text = text), options)
  # see my-app.js function
  session$sendCustomMessage(
    type = "notification",
    message = jsonlite::toJSON(
      message,
      auto_unbox = TRUE,
      json_verbatim = TRUE
    )
  )

}
```

Pay attention to the `icon` element: we can't convert shiny tags to JSON, so they
must be converted to character first. If you multiple parameters may contain tags,
you must treat them accordingly!

## Initialize the widget
On the JS side, we create an array containing all compatible widget names in two categories:

```javascript
const uiWidgets = ["gauge", "swiper", "searchbar"];
const serverWidgets = ["toast", "photoBrowser", "notification"];
const widgets = uiWidgets.concat(serverWidgets);
```

We then define the `activateWidget` function, only considering UI widgets. This function
takes the current widget as parameter and checks whether it is part of the `uiWidgets` array:

```javascript
if (uiWidgets.indexOf(widget) > -1) {
  // Do things
}
```

As there may be multiple widgets of the same time to initialize, we must loop through all
possible elements. This is where the class is important and must match the widget name.
For instance, the gauge has the `gauge` class and the methods are always `app.gauge.`.
How do we loop through multiple widgets? We use the jQuery `each` method:

```javascript
if (uiWidgets.indexOf(widget) > -1) {
  $("." + widget).each(function() {
    // Do things
  }
}
```

We see that `$("." + widget)` gives `$(".gauge)` when the widget is a gauge, which
will target all gauges one by one. Then for each gauge, we extract the configuration containing
all options passed by the end user. Remember that each element has a unique id. 
We extract the current element `$(this)` in the `$el` variable and search for a 
script tag pointing to the unique tag having `$el.attr("id")` as id. The configuration is parsed
to convert it to an object. Note that most of the time, Framework7 expects to have a `el`
attributes which simply contains the CSS selector of the current element, 
in other words its unique id `'#' + $el.attr("id")`:

```javascript
if (uiWidgets.indexOf(widget) > -1) {
  $("." + widget).each(function() {
    var $el = $(this);
    var config = $(document).find(
      "script[data-for='" + $el.attr("id") + "']"
    );
    config = JSON.parse(config.html());
    // add the id
    config.el = '#' + $el.attr("id");
  }
}
```

The final step is to initialize the widget, which is quite straightforward if we notice that
`app.gauge` is the same as `app["gauge"]`. We obtain the general code:

```javascript
app[widget].create(config);
```

For the server widgets, it is even simpler. We recover the message with a `Shiny.addCustomMessageHandler("type", callback)` and
initialize it. The only possible source of problem is the custom message `type` that must be the same 
as the one specified in the R function with `session$sendCustomMessage("type", message)`. We create
an `else` statement following the `if` condition:

```javascript
Shiny.addCustomMessageHandler(widget, function(message) {
  app[widget].create(message).open();
});
```

As shown in the above code, we can also immediately open the widget by chaining methods. 

The full JavaScript code may be found below.

```javascript
// Instantiate a widget
activateWidget = function(widget) {
  // Handle ui side widgets
  if (uiWidgets.indexOf(widget) > -1) {
    $("." + widget).each(function() {
      var $el = $(this);
      var config = $(document).find(
        "script[data-for='" + $el.attr("id") + "']"
      );
      config = JSON.parse(config.html());
      // add the id
      config.el = '#' + $el.attr("id");

      // feed the create method
      app[widget].create(config);
    });
  } else {
    // This concerns toasts, notifications, photoBrowser, ...
    // that don't have any UI element in the DOM before creating
    // the widget instance.
    Shiny.addCustomMessageHandler(widget, function(message) {
      app[widget].create(message).open();
    });
  }
};
```

The final step is to activate all widgets. We proceed with a loop:

```javascript
// Loop over all widgets to activate them
widgets.forEach(function(w) {
  activateWidget(w);
});
```

## Update widgets
We would like to develop a similar generalized interface to update any element in the DOM.
Instead of having `update_f7_gauge`, `update_f7_notification`, ... we want an `update_f7_instance` function. 

