# Mastering Shiny's events {#shiny-input-gems}
We've already seen a couple of Shiny JS __events__ since the beginning of this book. Advanced readers
probably know the `shiny:connected`, meaning that the client and server are properly initialized and all
internal methods/functions are available to the programmer. Below, we add more elements to the list, trying to
give practical examples. 

## Get the last changed input

### Motivations

We probably all had this question one day: How can I get the __last changed__ input in a Shiny app? There are already some methods like this [one](https://stackoverflow.com/questions/31250587/creating-shiny-reactive-variable-that-indicates-which-widget-was-last-modified): 

```{r, eval = FALSE}
runApp(
  shinyApp(
    ui = shinyUI(
      fluidPage(
        textInput('txt_a', 'Input Text A'),
        textInput('txt_b', 'Input Text B'),
        uiOutput('txt_c_out'),
        verbatimTextOutput("show_last")
      )
    ),
    server = function(input, output, session) {
      output$txt_c_out <- renderUI({
        textInput('txt_c', 'Input Text C')
      })
      
      values <- reactiveValues(
        lastUpdated = NULL
      )
      
      observe({
        lapply(names(input), function(x) {
          observe({
            input[[x]]
            values$lastUpdated <- x
          })
        })
      })
      
      output$show_last <- renderPrint({
        values$lastUpdated
      })
    }
  )
)
```

Shouldn't this be easier? Could we do that from the client instead, thereby reducing the server load?

### Invoke JS events
`shiny:inputchanged` is the event we are looking for. It is fired each time an input gets a new value. The related events has five properties:

  - __name__, the event name.
  - __value__, the new value.
  - __inputType__, the input type.
  - __binding__, the related input binding.
  - __el__ the related input DOM element.

You may try below:

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  tags$script(
  "$(document).on('shiny:inputchanged', function(event) {
    console.log(event);
  });"
  ), 
  textInput("test", "Test")
)

server <- function(input, output) {}

shinyApp(ui, server)
```

Changing the `textInput()` value fires the event. Contrary to what is mentioned in the online [documentation](https://shiny.rstudio.com/articles/js-events.html), __inputType__ does not always have a value. In this case, an alternative, is to access the related input binding and extract its name, as shown below:

```js
$(document).on('shiny:inputchanged', function(event) {
  Shiny.setInputValue(
    'pleaseStayHome', 
    {
      name: event.name, 
      value: event.value, 
      type: event.binding.name.split('.')[1]
    }
  );
});
```

If you use this code in a custom shiny template, it is possible that input bindings doesn't have name, which would thereby make `event.binding.name.split('.')[1]` crash, event.binding` being undefined.

::: {.warningblock data-latex=""}
For the `textInput()`, the event is also fired when moving the mouse cursor with the keyboard arrows, which is a sort of FALSE positive, since the value isn't changed. However, as `Shiny.setInputValue` only sets a new value when the input value really changed (unless the __priority__ is set to __event__), we avoids this edge case. As an exercice, you may try to add `{priority: 'event'}` to the above code. 
:::

`$(document).on('shiny:inputchanged')` is also cancellable, that is we may definitely prevent the input to change it's value, calling `event.preventDefault();`.

```{r, eval=FALSE}
ui <- fluidPage(
  tags$script(
  "$(document).on('shiny:inputchanged', function(event) {
    event.preventDefault();
  });"
  ), 
  textInput("test", "Test"),
  textOutput("val")
)

server <- function(input, output) {
  output$val <- renderText(input$test)
}

shinyApp(ui, server)
```

### Example

`{shinyMobile}` natively implements this feature that may be accessed with `input$lastInputChanged`.

```{r, eval = FALSE}
library(shinyMobile)
shinyApp(
  ui = f7Page(
    title = "My app",
    f7SingleLayout(
      navbar = f7Navbar(
        title = "Single Layout",
        hairline = FALSE,
        shadow = TRUE
      ),
      toolbar = f7Toolbar(
        position = "bottom",
        f7Link(label = "Link 1", href = "https://www.google.com"),
        f7Link(label = "Link 2", href = "https://www.google.com")
      ),
      # main content,
      f7Card(
        f7Text(inputId = "text", label = "Text"),
        f7Slider(
          inputId = "range1", 
          label = "Range", 
          min = 0, max = 2, 
          value = 1, 
          step = 0.1
        ),
        f7Stepper(
          inputId = "stepper1", 
          label = "Stepper", 
          min = 0, 
          max = 10, 
          value = 5
        ),
        verbatimTextOutput("lastChanged")
      )
    )
  ),
  server = function(input, output) {
    output$lastChanged <- renderPrint(input$lastInputChanged)
  }
)
```

This approach has the advantage not to overload the server part with complex logic.

### About `{shinylogs}`
The `{shinylogs}` [@R-shinylogs] package developed by [dreamRs](https://github.com/dreamRs/shinylogs) provide this feature with much more advanced options.

```{r, eval = FALSE}
library(shinylogs)

shinyApp(
  ui = fluidPage(
    numericInput("n", "n", 1),
    sliderInput("s", "s", min = 0, max = 10, value = 5),
    verbatimTextOutput("lastChanged")
  ),
  server = function(input, output, session) {
    # specific to shinylogs
    track_usage(storage_mode = store_null())
    output$lastChanged <- renderPrint({
      input$`.shinylogs_lastInput`
    })
  }
)
```

## Custom overlay screens
If you ever designed corporate production apps, you probably faced this situation where clients wanted
a loading screen, whenever a computation occurs or at startup. 
To date, one of the most comprehensive alternative is the `{waiter}` [@R-waiter] package by John Coene. 
It provide myriad of options to significantly enhance the perceived performance of your app. 
In the following, we'll focus on the `waiter_preloader()` and `waiter_on_busy()` functions. How does this work?

### Preloader
Under the hood, this feature relies on the `shiny:idle` event. At app startup, `shiny:idle` is triggered just after `shiny:connected` and `shiny:sessioninitialized`. Then, `shiny:idle` is also called each time a computation cycle is finished. 

Whenever we call `waiter_preloader()`, an HTML overlay is added in the DOM. Moreover, this extra JS code ensures to
hide the waiter when shiny is ready:

```js
window.ran = false;
$(document).on('shiny:idle', function(event){
  if(!window.ran)
    hide_waiter(id = null);
  window.ran = true;
});
```

As a security, `window.ran` avoids to run this code twice. As an example, consider this app with
a slider input and a plot output. We simulated a delay of three seconds to produce the plot:

```{r, eval=FALSE}
library(shiny)
library(waiter)

ui <- fluidPage(
  use_waiter(), # dependencies
  # shows before anything else 
  waiter_preloader(spin_fading_circles()), 
  sliderInput("obs", "Number of observations:",
    min = 0, max = 1000, value = 500
  ),
  plotOutput("distPlot")
)

server <- function(input, output){
  output$distPlot <- renderPlot({
    Sys.sleep(3)
    hist(rnorm(input$obs))
  })
}
shinyApp(ui, server)
```

Notice how the waiter correctly handles the plot processing time.

### Load on busy
Similarly, the `waiter_on_busy()` exploit the `shiny:idle` and `shiny:busy` events. We first create the loader as soon as 
shiny is busy:

```js
$(document).on('shiny:busy', function(event) {
  show_waiter(
    id = null,
    html = ..., 
    color = ...
  );
});
```

and is hidden once shiny is done:

```js
$(document).on('shiny:idle', function(event) {
  hide_waiter(null);
});
```