# Hidden gems about inputs {#shiny-input-gems}
We present some tools that may be useful...TO FINISH

## Get the last changed input

### Motivations

We probably all had this question one day: How can I get the last changed input in Shiny? There are already some methods like this [one](https://stackoverflow.com/questions/31250587/creating-shiny-reactive-variable-that-indicates-which-widget-was-last-modified) provided by Dean Attali. 

```{r, eval = FALSE}
runApp(
  shinyApp(
    ui = shinyUI(
      fluidPage(
        textInput('txt_a', 'Input Text A'),
        textInput('txt_b', 'Input Text B'),
        uiOutput('txt_c_out'),
        verbatimTextOutput("show_last")
      )
    ),
    server = function(input, output, session) {
      output$txt_c_out <- renderUI({
        textInput('txt_c', 'Input Text C')
      })
      
      values <- reactiveValues(
        lastUpdated = NULL
      )
      
      observe({
        lapply(names(input), function(x) {
          observe({
            input[[x]]
            values$lastUpdated <- x
          })
        })
      })
      
      output$show_last <- renderPrint({
        values$lastUpdated
      })
    }
  )
)
```

Shouldn't this be easier? Could we do that from the client instead, thereby reducing the server load?

### JavaScript, my friend

Here comes our friend, JavaScript. If you insert this snippet in the head of your app, you will be able to get the last changed input (name, value and type). 

```js
$(document).on('shiny:inputchanged', function(event) {
  Shiny.setInputValue('pleaseStayHome', {name: event.name, value: event.value, type: event.binding.name.split('.')[1]});
});
```

If you use this code in a custom shiny template, it is possible that input bindings don't have name, which would thereby make `event.binding.name.split('.')[1]` crash because `event.binding` is undefined. If so, you may remove this part:

```js
$(document).on('shiny:inputchanged', function(event) {
  Shiny.setInputValue('pleaseStayHome', {name: event.name, value: event.value});
});
```


### Example

`{shinyMobile}` natively implements this feature that may be accessed with `input$lastInputChanged`.

```{r, eval = FALSE}
library(shinyMobile)
shinyApp(
  ui = f7Page(
    title = "My app",
    f7SingleLayout(
      navbar = f7Navbar(
        title = "Single Layout",
        hairline = FALSE,
        shadow = TRUE
      ),
      toolbar = f7Toolbar(
        position = "bottom",
        f7Link(label = "Link 1", src = "https://www.google.com"),
        f7Link(label = "Link 2", src = "https://www.google.com", external = TRUE)
      ),
      # main content,
      f7Card(
        f7Text(inputId = "text", label = "Text"),
        f7Slider(inputId = "range1", label = "Range", min = 0, max = 2, value = 1, step = 0.1),
        f7Stepper(inputId = "stepper1", label = "Stepper", min = 0, max = 10, value = 5),
        verbatimTextOutput("lastChanged")
      )
    )
  ),
  server = function(input, output) {
    output$lastChanged <- renderPrint(input$lastInputChanged)
  }
)
```

This approach has the advantage not to overload the server part with complex logic.

### About {shinylogs}
The `{shinylogs}` package developed by [dreamRs](https://github.com/dreamRs/shinylogs) contains this feature with much more advanced options.

```{r, eval = FALSE}
library(shinylogs)

shinyApp(
  ui = fluidPage(
    numericInput("n", "n", 1),
    sliderInput("s", "s", min = 0, max = 10, value = 5),
    verbatimTextOutput("lastChanged")
  ),
  server = function(input, output, session) {
    # specific to shinylogs
    track_usage(storage_mode = store_null())
    output$lastChanged <- renderPrint(input$`.shinylogs_lastInput`)
  }
)
```